---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
Sys.setenv(LANGUAGE="en")
```

## Purpose of the package

The goal of `shapleyDecompo` is to compute the Shapley decomposition of the inequality of an outcome variable across different attributes from an econometric model.

## Installation instructions

You can install the development version of `shapleyDecompo` from [GitHub](https://github.com/kevinomics/shapleyDecompo) with:

``` r
# install.packages("devtools")
devtools::install_github("kevinomics/shapleyDecompo")
```

## Overview

The package can be used to compute the Shapley decomposition from a GLM (see `?stats::glm()`), or a Tobit II model obtained with the `sampleSelection` package (see `?sampleSelection::selection()`). The package provides with a function called `shapleyDecompo()` which can be used to compute the Shapley decomposition.

```{r load, message=FALSE}
library("shapleyDecompo")
```

```{r example}
data(Mroz87)
exTobitModel <- sampleSelection::selection(
  lfp ~ age + I(age^2) + faminc + kids5 + educ,
  wage ~ exper + I(exper^2) + educ + city,
  data = Mroz87)
shapleyDecompo(database = Mroz87,
               model_eco = exTobitModel,
               equation = "outcome",
               equaGame = FALSE,
               correction = NA,
               data_weights = rep(1, nrow(Mroz87)),
               residuals = FALSE,
               transfo = NULL,
               measure = Gini_w,
               theta = NULL)
```

The function `shapleyDecompo()` is a wrap function of several functions used to compute the decomposition:

1. The function `getFactorList()` is used to extract the list of factors in the model.
```{r getFactorList}
factors <- getFactorList(
   equation = "outcome",
   model_eco = exTobitModel,
   database = Mroz87,
   residuals = FALSE)
factors
```

2. The function `getCoalitions()` is used to obtain all possible coalitions from a list of factors.
```{r getCoalitions}
coa <- getCoalitions(factors_list = factors)
coa
```

3. The function `getShapleyDistrib()` is used to obtain the model matrix needed to perform the decomposition.
```{r getShapleyDistrib}
distrib <- getShapleyDistrib(model_eco = exTobitModel,
                             database = Mroz87,
                             equation = "outcome")
```

4. The function `getInequality()` is used to compute the inequality for a given coalition. Several functions in the package can be used to compute corresponding inequality measures (see `?Atkinson()`, `?Gini_G()`, `?Gini_w()`, `?Var()`, `?Entropy()`, `?Kolm()`).
```{r getInequality}
getInequality(coalition = coa[1, ],
              factors_list = factors,
              model_eco = exTobitModel,
              equaGame = FALSE,
              errors = NA,
              equation = "outcome",
              measure = Gini_w,
              database = Mroz87,
              mXOutcome = distrib,
              transfo = NULL,
              correction = NA,
              theta = NULL,
              weights = rep(1, nrow(Mroz87)))
```
The function `getInequality()` can be used in a loop to compute the inequalities of all coalitions.

```{r getInequalitLoop}
ineq <- vector()
for(i in 1:nrow(coa)){
  ineq[i] <- getInequality(coalition = coa[i, ],
              factors_list = factors,
              model_eco = exTobitModel,
              equaGame = FALSE,
              errors = NA,
              equation = "outcome",
              measure = Gini_w,
              database = Mroz87,
              mXOutcome = distrib,
              transfo = NULL,
              correction = NA,
              theta = NULL,
              weights = rep(1, nrow(Mroz87)))
}
ineq
```
5. The function `getMarginalContrib()` can be used to compute the marginal contributions of each attribute in all coalitions.

```{r getMarginalContrib}
margContrib <- getMarginalContrib(inequality = ineq,
                   coalitions = coa,
                   nVar = length(factors),
                   wMC = TRUE)
margContrib
```
6. The function `getShapleyVal()` can be used to compute the Shapley values from marginal contributions of attributes.
```{r getShapleyVal}
getShapleyVal(marginalContrib = margContrib)
```
7. The function `getInteractionTerms()` can be used to extract the interaction terms from the marginal contributions of attributes in each coalition.
```{r getInteractionTerms}
getInteractionTerms(factors_list = factors,
                    coalitions = coa,
                    marginalContrib = margContrib)
```


## Citation

To cite this package please use: 

```{r auth01, eval=TRUE}
citation("shapleyDecompo") 
```

and/or: Fourrey, K. (2023). A Regression-Based Shapley Decomposition for Inequality Measures. Annals of Economics and Statistics, 149, 39–62. https://doi.org/10.2307/48718079
