---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
Sys.setenv(LANGUAGE="en")
```

# Purpose of the package

The goal of `shapleyDecompo` is to compute the Shapley decomposition of the inequality of an outcome variable across different attributes from an econometric model.

## Installation instructions

You can install the development version of shapleyDecompo from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("kevinomics/shapleyDecompo")
```

## Overview

The package can be used to compute the Shapley decomposition from a GLM (see `?stats::glm()`), or a Tobit II model obtained with the `sampleSelection` package (see `?sampleSelection::selection()`). The package provides with a function called `shapleyDecompo()` which can be used to compute the Shapley decomposition.

```{r load, message=FALSE}
library("shapleyDecompo")
```

```{r example}
shapleyDecompo(database = exData,
               model_eco = exTobitModel,
               equation = "outcome",
               equaGame = TRUE,
               correction = NA,
               data_weights = exData$extridf,
               residuals = TRUE,
               transfo = exp,
               measure = Atkinson,
               theta = 1)
```

## Citation

To cite this package please use: 

```{r auth01, eval=TRUE}
citation("shapleyDecompo") 
```

and/or: Fourrey, K. (2023). A Regression-Based Shapley Decomposition for Inequality Measures. Annals of Economics and Statistics, 149, 39–62. https://doi.org/10.2307/48718079
